### 2022-03-11 paper/paper.tex
Add ~ to citations and references

Perhaps the best motivation for ``end-to-end'' classification comes from showing, as in Figure \ref{fig:traditional-pipeline}, what end-to-end classification is \textit{not}.
In particular, the steps ``Particle flow reconstruction'' and ``Feature engineering'', outlined in \myhref{ss:three-methods}{Section \ref{ss:three-methods}}, turn out to be quite non-trivial.
End-to-end classification removes these steps.
In an end-to-end workflow, shown schematically in Figure \ref{fig:cnn-pipeline}, a classifier outputs a result computed directly from raw detector data, with minimal intermediate data processing.

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.70\linewidth]{vector/cnn-pipeline.pdf}
    \caption{An end-to-end classification workflow eliminates the complicated intermediate steps involved in traditional classification (cf. Figure \ref{fig:traditional-pipeline}).}
    \label{fig:cnn-pipeline}
\end{figure}

 As implied by Figures \ref{fig:traditional-pipeline} and \ref{fig:cnn-pipeline}, particle classification requires as input the data measured by a particle detector.
 In the next section, we explore the nature of this detector data, and outline how the data is produced and measured at the Large Hadron Collider.

\section{How the LHC Produces and Measures Collision Data}
For our purposes, the Large Hadron Collider (LHC) is a synchrotron accelerator that collides protons at teravolt-order center-of-mass energy \cite{lhc}.
Each collision produces a cascade of secondary particles, whose energies and trajectories are measured in specialized detectors like CMS and ATLAS.
The purpose of this section is to explain:
\begin{enumerate}

@@

    % \item how the pixels and colors in Figure \ref{fig:data-raw-ecal} correspond directly to the energies and trajectories of physical particles in an actual collision event.

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.75\linewidth]{vector/cern-complex.pdf}
    \caption{The components of the CERN accelerator complex relevant to this work.
    Adapted from \cite{image-cern-complex}.}
    \label{fig:cern-complex}
\end{figure}

The story begins with a bottle of hydrogen gas.
A strong electric field strips the hydrogen atoms of their electrons, leaving behind only protons.
These protons are accelerated through the linear accelerator \textit{Linac2},\footnote{In the LHC's 2020 high luminosity upgrade, Linac2 is replaced with Linac4, which accelerates negative hydrogen ions instead \cite{linac4-tdr}.} setting in motion a journey through the maze of synchrotron boosting stages shown in Figure \ref{fig:cern-complex} and culminating at the LHC.
Depending on operating conditions, the LHC accelerates protons to an ultimate kinetic energy of up to \SI{6.5}{\tera \electronvolt}; these protons travel around the LHC in two opposing beamlines in bunches of roughly $ 10^{11} $ particles, separated in time by \SI{25}{\nano \second}.
The beamlines cross, allowing for proton-proton collisions, at four nominal collision points, around which are centered, in anticipation, the LHC's four main detectors: ALICE, CMS, LHCb, and ATLAS.

@@

\end{enumerate}
Trajectories are measured with instruments called \textit{trackers}, while energy is measured with \textit{calorimeters}.
From trajectory and energy, and by noting which particles appear in which subdetector, we can reconstruct a number of secondary quantities, including particle momentum, identity, and the positions of production and decay vertices.
In this section we first define the detector coordinate system, then discuss each measurement instrument in turn.
For concreteness, we will focus on the instruments used at the Compact Muon Solenoid (CMS) detector \cite{cms}, but similar concepts apply to most particle detectors currently used in high-energy physics.
For better orientation, each of the four CMS sub-detectors discussed in this section appear, in correct scale, in the diagram of the CMS detector shown in Figure \ref{fig:cms} two pages below.

\subsubsection{The Detector Coordinate System at the CMS}
The CMS coordinate system is best understood visually, as shown in Figure \ref{fig:cms-coordinate-system}.
The detector uses a hybrid Cartesian/cylindrical coordinate system, in which the origin coincides with the nominal collision point, while the $ x $, $ y $, and $ z $ axes point towards the center of the LHC, vertically upward, and along the beamline, respectively.
The azimuthal angle $ \varphi $ is measured in the $ xy $ plane, while the polar angle $ \theta $ is measured in the $ yz $ plane.
Conventionally, the polar angle is given in terms of a quantity called \textit{pseudorapidity},\footnote{Particle physicists work with pseudorapidity, and a related quantity called rapidity, because differences in these quantities are Lorentz-invariant to boosts along the beam axis.
For our purposes, it is enough to view pseudorapidity as just an alternate representation of the polar angle $ \theta $.} denoted by $ \eta $ and defined as
\begin{equation}
    \eta \equiv - \ln \left( \tan \frac{\theta}{2} \right) \implies \theta = 2 \arctan e^{- \eta}. \label{eq:pseudorapidity}
\end{equation}

@@

The CMS tracker \cite{tracker-tdr} consists of millions of these silicon elements arranged in concentric, cylindrical layers around the nominal collision point.
An incident particle passing through a silicon pixel (strip) excites the atoms in the pixel's (strip's) depletion region, creating electron-hole pairs.
The silicon element's electric field then accelerates the electrons and holes to opposite pixel (strip) faces, leading to a measurable charge pulse called a \textit{hit}.
Hits from successive layers are combined to reconstruct an incident particle's trajectory.

@@

\subsubsection{Electromagnetic Calorimeters}
Electromagnetic calorimeters (ECALs) measure the energy of particles that interact with matter via the electromagnetic interaction.
At the CMS, the basic ECAL building block is a lead tungstate (\chem{PbWO_4}) scintillator crystal connected to a photodetector \cite{ecal-tdr}.

The basic working principle is as follows: high-energy incident particles interact with the \chem{PbWO_4} scintillators (electrons via bremsstrahlung, high-energy photons predominantly via pair production) to produce secondary \textit{electromagnetic showers}---cascades of lower-energy photons, electrons, and positrons.
Shower particles excite the \chem{PbWO_4} scintillators, which emit blue-light \textit{scintillation photons} during the relaxation process.
These scintillation photons reach the crystal's attached photodetector (avalanche photodiodes in the ECAL barrel and vacuum phototriodes in the ECAL end caps \cite{ecal-tdr}), where they free electrons via the photoelectric effect.
After further amplification, these photoelectrons produce a measurable electric current whose amplitude encodes the energy deposited by the initial incident particle.

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.9\linewidth]{raster/png/cms.png}
    \caption{A cutaway diagram of the CMS detector, showing each of the sub-detectors (silicon trackers, ECAL, HCAL, and muon detectors) discussed in this work.
    \cite{image-cms}}
    \label{fig:cms}
\end{figure}

\subsubsection{Hadronic Calorimeters}
Hadronic calorimeters (HCALs) measure the energy of particles that interact with matter via the strong interaction.
The CMS HCAL consists of alternating layers of absorbers---brass and steel plates---and scintillation detection modules---plastic scintillator tiles connected to hybrid photodiodes \cite{hcal-tdr}.

@@

\subsubsection{Muon Detection}
For the purposes of this work, this section may be skipped without loss of continuity.
However, for completeness, we now briefly outline the CMS's namesake muon detection system.
The CMS muon detector is a gas-based detector, whose basic module is a rectangular \textit{drift tube} enclosing a mixture of argon and \chem{CO_2} gas (10 to 20 percent \chem{CO_2} by volume).
Through the chamber's center runs a thin steel anode wire (radius \SI{50}{\micro \meter}), held at a kilovolt-order potential difference relative to copper-coated electrodes glued to the chamber's outer surface \cite{muon-tdr}.

@@

Three other muon detector components, namely resistive plate chambers, cathode strip chambers, and gas electron multipliers, rely on the same physical principle of gas ionization and ion collection.
For more information, interested readers are referred to Refs. \cite{muon-web} or \cite{muon-tdr} for a friendly or technical discussion, respectively.

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=\linewidth]{raster/raster-svg/event-image.pdf}
    \caption{A composite image of a photon+jet (background) event at the CMS.
    Tracks appear in orange log scale, ECAL hits in blue log scale, and HCAL hits in gray linear scale.
    Pixels correspond to physical \chem{PbWO_4} scintillator crystals in the ECAL barrel's $ 360 \times 170 $ crystal grid, which spans $ \varphi \in [0, 2\pi) $ and $ \abs{\eta} < 1.479$.
    Adapted from \cite{andrews-higgs}.}
    \label{fig:data-raw-ecal}
\end{figure}

If we combine the Tracker hits and ECAL/HCAL signal pulses and project these quantities onto the detector's $ \varphi \eta $ plane, we get something like the image shown in Figure \ref{fig:data-raw-ecal}.
This image, from a simulated photon+jet event at the CMS \cite{data-bg-jet}, shows the raw Tracker (position) and ECAL/HCAL (energy) information encoding the event, projected onto the ECAL barrel's $ 360 \times 170 $ grid of \chem{PbWO_4} scintillator crystals.

Figure \ref{fig:data-raw-ecal} serves as our basic model of low-level detector data: an image-like grid with two spatial dimensions $ (\varphi, \eta) $ and three detector channels, in which pixel intensity corresponds to either (i) charge deposited in the Tracker, or (ii) energy deposited in ECAL or HCAL.
In each case, the position of a given pixel in the image has a well-defined physical correspondence to the position of a measurement module (e.g. Tracker pixel, ECAL scintillation crystal) in the detector.
Equipped with this form of image-based data, we have three options:
\begin{enumerate}[(a)]

@@

    \item \label{item:kinematic-classification} \textit{Kinematic-based classification}, in which we feed the raw data into a \textit{particle flow reconstruction} algorithm,\footnote{Particle flow reconstruction falls largely beyond the scope of this work.
    As an example, we can reconstruct a particle's transverse momentum from the curvature of its trajectory in the detector's magnetic field, but we will otherwise treat reconstruction as a ``black box'' that outputs kinematic quantities describing a collision.
    Interested readers are referred to Ref. \cite{particle-flow}.} which reconstructs \textit{kinematic features}, for example transverse momentum (relative to the beam axis) and pseudorapidity (Eq. \ref{eq:pseudorapidity}), describing the decay products.
    We then use this kinematic data as input into a fully-connected neural network (FCN), which outputs a classification result.

    \item \label{item:high-level-classification} Reconstruct kinematic features from the low-level data as above, then use these kinematic features to hand-engineer \textit{high-level features}, manually designed to separate signal from background events on an experiment-by-experiment basis (for example, decay particle invariant masses at which we expect high signal production).
    Then, feed these high-level features, often in combination with the kinematic features described above, into a FCN or BDT classifier.
    See Ref. \cite{baldi-higgs} for a study comparing kinematic-based and high-level classification.% , which outputs classification results.

@@

Schematically, classification workflows \ref{item:kinematic-classification} and \ref{item:high-level-classification} resemble Figure \ref{fig:fcn-in-out}.
\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.9\linewidth]{vector/fcn-in-out.pdf}
    \caption{The binary classification process using a FCN.
    The outputs $ \hat{y}_{\text{sig}} $ and $ \hat{y}_{\text{bg}} $ represent the predicted probabilities that the inputted feature vector $ \vec{x}_{\text{feature}} $ corresponds to either signal or background, respectively, as explained in \myhref{ss:output}{Section \ref{ss:output}}.}
    \label{fig:fcn-in-out}
\end{figure}

@@

\section{Fully-Connected Networks} \label{s:fcn}
In this section we take a look inside the ``Fully-Connected Neural Network'' box in Figure \ref{fig:fcn-in-out}.
Suppose, as in Figure \ref{fig:fcn-in-out}, that we have a set of kinematic properties encoding a collision event (found, for example, with the particle flow reconstruction algorithm mentioned in \myhref{ss:three-methods}{Section \ref{ss:three-methods}}), which we pack into a vector $ \vec{x} $.
In machine learning terminology \cite{homl}, each individual kinematic property is called a \textit{feature}, and the two possible outcomes (signal or background) are called \textit{classes}.
The correct class (either signal or background) for a given event is called the event's \textit{target} or \textit{label}, and the complete set of features describing a single event, together with the label, is called an \textit{instance}.
Finally, the complete set of available instances (the complete set of collision events) forms a \textit{dataset}.
With the vocabulary lesson out of the way, we are equipped to formally describe a neural network.
But first, we should reveal supervised machine learning's dirty secret:
\begin{quote}
    The dataset used to train a neural network is entirely \textit{simulated}.\footnote{Simulated collision data is constructed with sophisticated algorithms involving Monte-Carlo sampling, a theoretical physical framework such as the Standard Model, and a geometrical model of the particle detector.
    Like particle flow reconstruction, this falls beyond the scope of this work.}
\end{quote}
Why? Well, we \textit{know} the outcome of each simulated collision event.
Thus, our dataset comes with correct labels for every single instance.
Using this large dataset of correctly-labeled data, we can train a neural network to give accurate results even when, as in practical applications, the correct results aren't known beforehand.

@@

We now pause to clarify what a binary classifier's output looks like and how to interpret the result.
First, we decide on an order in which to quote the possible classes: in this work we will quote signal first, then background.
The label, represented by a vector $ \vec{y} $, equals one at the position of the correct class and zero at the position of the incorrect class.
In the signal-first convention, a generic label vector reads $ \vec{y} = (y_{\text{sig}}, y_{\text{bg}}) $; the label $ \vec{y} = (1, 0) $ represents signal while $ \vec{y} = (0, 1) $ represents background.
Meanwhile, a classifier's prediction, denoted by $ \hat{\y} = (\hat{y}_{\text{sig}}, \hat{y}_{\text{bg}})$, represents the predicted \textit{probabilities} that an event corresponds to either signal or background; these two values should sum to one.
Figure \ref{fig:output} summarizes these ideas visually.

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.95\linewidth]{vector/fcn-architecture.pdf}
    \caption{A fully-connected network's internal architecture; each neuron beyond the input layer connects to \textit{every} neuron in the previous layer.
    The figure uses the notation from \myhref{s:appendix-fcn}{Appendix \ref{s:appendix-fcn}}.
    The $ \sum $ (sum) symbol represents the multiply-add neuron operation in Equation \ref{eq:preactivation}, while the hinge and $ \sigma $ symbols represent the ReLU and softmax activation functions, shown in Figure \ref{fig:relu} and Equation \ref{eq:softmax}, respectively.}
    \label{fig:fcn-architecture}
\end{figure}

For better orientation, Figure \ref{fig:fcn-architecture} shows the architecture of a fully-connected neural network.
The basic building of a FCN is a \textit{neuron}---represented by the circular nodes in Figure \ref{fig:fcn-architecture}.
Neurons are then grouped into one-dimensional stacks to form \textit{layers}.
A FCN consists of a single \textit{input layer}, an arbitrary number of \textit{hidden layers}, and a single \textit{output layer}, which outputs the classification scores discussed in Section \ref{ss:output}.

In this section we build up to a big-picture interpretation of FCN-based classification as the optimization of a vector-valued function, a concept which should be familiar to physicists.
That said, before proceeding we first refer interested readers to \myhref{s:appendix-fcn}{Appendix \ref{s:appendix-fcn}}, which thoroughly explains the nuts and bolts of a FCN's internal architecture in the language of the machine-learning literature.
However, while it might help demystify the ``black box'' nature of neural networks, this formal treatment can be skipped without loss of continuity.

@@

\end{enumerate}
As required by point \ref{item:neuron-input} and visualized in Figure \ref{fig:fcn-architecture}, each hidden-layer neuron has input connections to \textit{every} neuron in the previous layer, hence the name ``fully-connected'' network.
The strength of each of these connections is parameterized by a unique scalar \textit{weight} $ w_{j} \in \mathbb{R} $, where $  j = 1, 2, \ldots, n_{\text{prev}} $; in other words, $ w_{j} $ parameterizes the current neuron's connection to the $ j $-th neuron in the previous layer.
For compactness, we write the current neuron's weights as the vector
\begin{equation*}
    \w = \big( w_{1}, w_{2}, \ldots, w_{n_{\text{prev}}} \big) \in \mathbb{R}^{n_{\text{prev}}};
\end{equation*}
similarly, the previous layer's activation values are written as the vector
\begin{equation*}
    \a_{\text{prev}} = \Big( a_{1}^{(\text{prev})}, a_{2}^{(\text{prev})}, \ldots, a_{n_{\text{prev}}}^{(\text{prev})} \Big) \in \mathbb{R}^{n_{\text{prev}}}.
\end{equation*}
In addition to its $ n_{\text{prev}} $ weights, the current neuron is parameterized by a single scalar bias $ b $, which serves as an additive constant which can translate the neuron's activation value.

@@

\begin{wrapfigure}{R}{0.36\textwidth}
    \centering
    \includegraphics[width=\linewidth]{vector/relu.pdf}
    \caption{The ReLU function}
    \label{fig:relu}
\end{wrapfigure}
Crucially, using \textit{non-linear} activation functions allows the network to form non-linear decision boundaries in the high-dimensional feature space of the inputted feature vectors; these non-linear boundaries enormously improve classification potential.
In practice, deep networks predominantly use the rectified linear unit (ReLU) function (Figure \ref{fig:relu}) or one of its variants;\footnote{See Wikipedia's \href{https://en.wikipedia.org/wiki/Activation_function\#Comparison_of_activation_functions}{\underline{list of common activation functions}} for a comprehensive list.} I will use the notation $ f_{\text{a}} $ for generality.

@@

A discrete convolution involves two multidimensional objects---an input image and a \textit{kernel} or \textit{filter}.
The kernel is a pixel-like grid like the input image, but always smaller in height and width.
Each kernel pixel is assigned a scalar \textit{weight}, and the kernel as a whole a single scalar \textit{bias}; these kernel weights and biases are a CNN's tunable parameters, and replace the neuron weights and biases in a FCN \cite{homl, goodfellow}.

>>>>

Perhaps the best motivation for ``end-to-end'' classification comes from showing, as in Figure~\ref{fig:traditional-pipeline}, what end-to-end classification is \textit{not}.
In particular, the steps ``Particle flow reconstruction'' and ``Feature engineering'', outlined in \myhref{ss:three-methods}{Section~\ref{ss:three-methods}}, turn out to be quite non-trivial.
End-to-end classification removes these steps.
In an end-to-end workflow, shown schematically in Figure~\ref{fig:cnn-pipeline}, a classifier outputs a result computed directly from raw detector data, with minimal intermediate data processing.

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.70\linewidth]{vector/cnn-pipeline.pdf}
    \caption{An end-to-end classification workflow eliminates the complicated intermediate steps involved in traditional classification (cf. Figure~\ref{fig:traditional-pipeline}).}
    \label{fig:cnn-pipeline}
\end{figure}

 As implied by Figures~\ref{fig:traditional-pipeline} and~\ref{fig:cnn-pipeline}, particle classification requires as input the data measured by a particle detector.
 In the next section, we explore the nature of this detector data, and outline how the data is produced and measured at the Large Hadron Collider.

\section{How the LHC Produces and Measures Collision Data}
For our purposes, the Large Hadron Collider (LHC) is a synchrotron accelerator that collides protons at teravolt-order center-of-mass energy~\cite{lhc}.
Each collision produces a cascade of secondary particles, whose energies and trajectories are measured in specialized detectors like CMS and ATLAS.
The purpose of this section is to explain:
\begin{enumerate}

@@

    % \item how the pixels and colors in Figure~\ref{fig:data-raw-ecal} correspond directly to the energies and trajectories of physical particles in an actual collision event.

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.75\linewidth]{vector/cern-complex.pdf}
    \caption{The components of the CERN accelerator complex relevant to this work.
    Adapted from~\cite{image-cern-complex}.}
    \label{fig:cern-complex}
\end{figure}

The story begins with a bottle of hydrogen gas.
A strong electric field strips the hydrogen atoms of their electrons, leaving behind only protons.
These protons are accelerated through the linear accelerator \textit{Linac2},\footnote{In the LHC's 2020 high luminosity upgrade, Linac2 is replaced with Linac4, which accelerates negative hydrogen ions instead~\cite{linac4-tdr}.} setting in motion a journey through the maze of synchrotron boosting stages shown in Figure~\ref{fig:cern-complex} and culminating at the LHC.
Depending on operating conditions, the LHC accelerates protons to an ultimate kinetic energy of up to \SI{6.5}{\tera \electronvolt}; these protons travel around the LHC in two opposing beamlines in bunches of roughly $ 10^{11} $ particles, separated in time by \SI{25}{\nano \second}.
The beamlines cross, allowing for proton-proton collisions, at four nominal collision points, around which are centered, in anticipation, the LHC's four main detectors: ALICE, CMS, LHCb, and ATLAS.

@@

\end{enumerate}
Trajectories are measured with instruments called \textit{trackers}, while energy is measured with \textit{calorimeters}.
From trajectory and energy, and by noting which particles appear in which subdetector, we can reconstruct a number of secondary quantities, including particle momentum, identity, and the positions of production and decay vertices.
In this section we first define the detector coordinate system, then discuss each measurement instrument in turn.
For concreteness, we will focus on the instruments used at the Compact Muon Solenoid (CMS) detector~\cite{cms}, but similar concepts apply to most particle detectors currently used in high-energy physics.
For better orientation, each of the four CMS sub-detectors discussed in this section appear, in correct scale, in the diagram of the CMS detector shown in Figure~\ref{fig:cms} two pages below.

\subsubsection{The Detector Coordinate System at the CMS}
The CMS coordinate system is best understood visually, as shown in Figure~\ref{fig:cms-coordinate-system}.
The detector uses a hybrid Cartesian/cylindrical coordinate system, in which the origin coincides with the nominal collision point, while the $ x $, $ y $, and $ z $ axes point towards the center of the LHC, vertically upward, and along the beamline, respectively.
The azimuthal angle $ \varphi $ is measured in the $ xy $ plane, while the polar angle $ \theta $ is measured in the $ yz $ plane.
Conventionally, the polar angle is given in terms of a quantity called \textit{pseudorapidity},\footnote{Particle physicists work with pseudorapidity, and a related quantity called rapidity, because differences in these quantities are Lorentz-invariant to boosts along the beam axis.
For our purposes, it is enough to view pseudorapidity as just an alternate representation of the polar angle $ \theta $.} denoted by $ \eta $ and defined as
\begin{equation}
    \eta \equiv - \ln \left( \tan \frac{\theta}{2} \right) \implies \theta = 2 \arctan e^{- \eta}. \label{eq:pseudorapidity}
\end{equation}

@@

The CMS tracker~\cite{tracker-tdr} consists of millions of these silicon elements arranged in concentric, cylindrical layers around the nominal collision point.
An incident particle passing through a silicon pixel (strip) excites the atoms in the pixel's (strip's) depletion region, creating electron-hole pairs.
The silicon element's electric field then accelerates the electrons and holes to opposite pixel (strip) faces, leading to a measurable charge pulse called a \textit{hit}.
Hits from successive layers are combined to reconstruct an incident particle's trajectory.

@@

\subsubsection{Electromagnetic Calorimeters}
Electromagnetic calorimeters (ECALs) measure the energy of particles that interact with matter via the electromagnetic interaction.
At the CMS, the basic ECAL building block is a lead tungstate (\chem{PbWO_4}) scintillator crystal connected to a photodetector~\cite{ecal-tdr}.

The basic working principle is as follows: high-energy incident particles interact with the \chem{PbWO_4} scintillators (electrons via bremsstrahlung, high-energy photons predominantly via pair production) to produce secondary \textit{electromagnetic showers}---cascades of lower-energy photons, electrons, and positrons.
Shower particles excite the \chem{PbWO_4} scintillators, which emit blue-light \textit{scintillation photons} during the relaxation process.
These scintillation photons reach the crystal's attached photodetector (avalanche photodiodes in the ECAL barrel and vacuum phototriodes in the ECAL end caps~\cite{ecal-tdr}), where they free electrons via the photoelectric effect.
After further amplification, these photoelectrons produce a measurable electric current whose amplitude encodes the energy deposited by the initial incident particle.

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.9\linewidth]{raster/png/cms.png}
    \caption{A cutaway diagram of the CMS detector, showing each of the sub-detectors (silicon trackers, ECAL, HCAL, and muon detectors) discussed in this work.
   ~\cite{image-cms}}
    \label{fig:cms}
\end{figure}

\subsubsection{Hadronic Calorimeters}
Hadronic calorimeters (HCALs) measure the energy of particles that interact with matter via the strong interaction.
The CMS HCAL consists of alternating layers of absorbers---brass and steel plates---and scintillation detection modules---plastic scintillator tiles connected to hybrid photodiodes~\cite{hcal-tdr}.

@@

\subsubsection{Muon Detection}
For the purposes of this work, this section may be skipped without loss of continuity.
However, for completeness, we now briefly outline the CMS's namesake muon detection system.
The CMS muon detector is a gas-based detector, whose basic module is a rectangular \textit{drift tube} enclosing a mixture of argon and \chem{CO_2} gas (10 to 20 percent \chem{CO_2} by volume).
Through the chamber's center runs a thin steel anode wire (radius \SI{50}{\micro \meter}), held at a kilovolt-order potential difference relative to copper-coated electrodes glued to the chamber's outer surface~\cite{muon-tdr}.

@@

Three other muon detector components, namely resistive plate chambers, cathode strip chambers, and gas electron multipliers, rely on the same physical principle of gas ionization and ion collection.
For more information, interested readers are referred to Refs.~\cite{muon-web} or~\cite{muon-tdr} for a friendly or technical discussion, respectively.

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=\linewidth]{raster/raster-svg/event-image.pdf}
    \caption{A composite image of a photon+jet (background) event at the CMS.
    Tracks appear in orange log scale, ECAL hits in blue log scale, and HCAL hits in gray linear scale.
    Pixels correspond to physical \chem{PbWO_4} scintillator crystals in the ECAL barrel's $ 360 \times 170 $ crystal grid, which spans $ \varphi \in [0, 2\pi) $ and $ \abs{\eta} < 1.479$.
    Adapted from~\cite{andrews-higgs}.}
    \label{fig:data-raw-ecal}
\end{figure}

If we combine the Tracker hits and ECAL/HCAL signal pulses and project these quantities onto the detector's $ \varphi \eta $ plane, we get something like the image shown in Figure~\ref{fig:data-raw-ecal}.
This image, from a simulated photon+jet event at the CMS~\cite{data-bg-jet}, shows the raw Tracker (position) and ECAL/HCAL (energy) information encoding the event, projected onto the ECAL barrel's $ 360 \times 170 $ grid of \chem{PbWO_4} scintillator crystals.

Figure~\ref{fig:data-raw-ecal} serves as our basic model of low-level detector data: an image-like grid with two spatial dimensions $ (\varphi, \eta) $ and three detector channels, in which pixel intensity corresponds to either (i) charge deposited in the Tracker, or (ii) energy deposited in ECAL or HCAL.
In each case, the position of a given pixel in the image has a well-defined physical correspondence to the position of a measurement module (e.g. Tracker pixel, ECAL scintillation crystal) in the detector.
Equipped with this form of image-based data, we have three options:
\begin{enumerate}[(a)]

@@

    \item \label{item:kinematic-classification} \textit{Kinematic-based classification}, in which we feed the raw data into a \textit{particle flow reconstruction} algorithm,\footnote{Particle flow reconstruction falls largely beyond the scope of this work.
    As an example, we can reconstruct a particle's transverse momentum from the curvature of its trajectory in the detector's magnetic field, but we will otherwise treat reconstruction as a ``black box'' that outputs kinematic quantities describing a collision.
    Interested readers are referred to Ref.~\cite{particle-flow}.} which reconstructs \textit{kinematic features}, for example transverse momentum (relative to the beam axis) and pseudorapidity (Eq.~\ref{eq:pseudorapidity}), describing the decay products.
    We then use this kinematic data as input into a fully-connected neural network (FCN), which outputs a classification result.

    \item \label{item:high-level-classification} Reconstruct kinematic features from the low-level data as above, then use these kinematic features to hand-engineer \textit{high-level features}, manually designed to separate signal from background events on an experiment-by-experiment basis (for example, decay particle invariant masses at which we expect high signal production).
    Then, feed these high-level features, often in combination with the kinematic features described above, into a FCN or BDT classifier.
    See Ref.~\cite{baldi-higgs} for a study comparing kinematic-based and high-level classification.% , which outputs classification results.

@@

Schematically, classification workflows~\ref{item:kinematic-classification} and~\ref{item:high-level-classification} resemble Figure~\ref{fig:fcn-in-out}.
\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.9\linewidth]{vector/fcn-in-out.pdf}
    \caption{The binary classification process using a FCN.
    The outputs $ \hat{y}_{\text{sig}} $ and $ \hat{y}_{\text{bg}} $ represent the predicted probabilities that the inputted feature vector $ \vec{x}_{\text{feature}} $ corresponds to either signal or background, respectively, as explained in \myhref{ss:output}{Section~\ref{ss:output}}.}
    \label{fig:fcn-in-out}
\end{figure}

@@

\section{Fully-Connected Networks} \label{s:fcn}
In this section we take a look inside the ``Fully-Connected Neural Network'' box in Figure~\ref{fig:fcn-in-out}.
Suppose, as in Figure~\ref{fig:fcn-in-out}, that we have a set of kinematic properties encoding a collision event (found, for example, with the particle flow reconstruction algorithm mentioned in \myhref{ss:three-methods}{Section~\ref{ss:three-methods}}), which we pack into a vector $ \vec{x} $.
In machine learning terminology~\cite{homl}, each individual kinematic property is called a \textit{feature}, and the two possible outcomes (signal or background) are called \textit{classes}.
The correct class (either signal or background) for a given event is called the event's \textit{target} or \textit{label}, and the complete set of features describing a single event, together with the label, is called an \textit{instance}.
Finally, the complete set of available instances (the complete set of collision events) forms a \textit{dataset}.
With the vocabulary lesson out of the way, we are equipped to formally describe a neural network.
But first, we should reveal supervised machine learning's dirty secret:
\begin{quote}
    The dataset used to train a neural network is entirely \textit{simulated}.\footnote{Simulated collision data is constructed with sophisticated algorithms involving Monte-Carlo sampling, a theoretical physical framework such as the Standard Model, and a geometrical model of the particle detector.
    Like particle flow reconstruction, this falls beyond the scope of this work.}
\end{quote}
Why? Well, we \textit{know} the outcome of each simulated collision event.
Thus, our dataset comes with correct labels for every single instance.
Using this large dataset of correctly-labeled data, we can train a neural network to give accurate results even when, as in practical applications, the correct results aren't known beforehand.

@@

We now pause to clarify what a binary classifier's output looks like and how to interpret the result.
First, we decide on an order in which to quote the possible classes: in this work we will quote signal first, then background.
The label, represented by a vector $ \vec{y} $, equals one at the position of the correct class and zero at the position of the incorrect class.
In the signal-first convention, a generic label vector reads $ \vec{y} = (y_{\text{sig}}, y_{\text{bg}}) $; the label $ \vec{y} = (1, 0) $ represents signal while $ \vec{y} = (0, 1) $ represents background.
Meanwhile, a classifier's prediction, denoted by $ \hat{\y} = (\hat{y}_{\text{sig}}, \hat{y}_{\text{bg}})$, represents the predicted \textit{probabilities} that an event corresponds to either signal or background; these two values should sum to one.
Figure~\ref{fig:output} summarizes these ideas visually.

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.95\linewidth]{vector/fcn-architecture.pdf}
    \caption{A fully-connected network's internal architecture; each neuron beyond the input layer connects to \textit{every} neuron in the previous layer.
    The figure uses the notation from \myhref{s:appendix-fcn}{Appendix~\ref{s:appendix-fcn}}.
    The $ \sum $ (sum) symbol represents the multiply-add neuron operation in Equation~\ref{eq:preactivation}, while the hinge and $ \sigma $ symbols represent the ReLU and softmax activation functions, shown in Figure~\ref{fig:relu} and Equation~\ref{eq:softmax}, respectively.}
    \label{fig:fcn-architecture}
\end{figure}

For better orientation, Figure~\ref{fig:fcn-architecture} shows the architecture of a fully-connected neural network.
The basic building of a FCN is a \textit{neuron}---represented by the circular nodes in Figure~\ref{fig:fcn-architecture}.
Neurons are then grouped into one-dimensional stacks to form \textit{layers}.
A FCN consists of a single \textit{input layer}, an arbitrary number of \textit{hidden layers}, and a single \textit{output layer}, which outputs the classification scores discussed in Section~\ref{ss:output}.

In this section we build up to a big-picture interpretation of FCN-based classification as the optimization of a vector-valued function, a concept which should be familiar to physicists.
That said, before proceeding we first refer interested readers to \myhref{s:appendix-fcn}{Appendix~\ref{s:appendix-fcn}}, which thoroughly explains the nuts and bolts of a FCN's internal architecture in the language of the machine-learning literature.
However, while it might help demystify the ``black box'' nature of neural networks, this formal treatment can be skipped without loss of continuity.

@@

\end{enumerate}
As required by point~\ref{item:neuron-input} and visualized in Figure~\ref{fig:fcn-architecture}, each hidden-layer neuron has input connections to \textit{every} neuron in the previous layer, hence the name ``fully-connected'' network.
The strength of each of these connections is parameterized by a unique scalar \textit{weight} $ w_{j} \in \mathbb{R} $, where $  j = 1, 2, \ldots, n_{\text{prev}} $; in other words, $ w_{j} $ parameterizes the current neuron's connection to the $ j $-th neuron in the previous layer.
For compactness, we write the current neuron's weights as the vector
\begin{equation*}
    \w = \big( w_{1}, w_{2}, \ldots, w_{n_{\text{prev}}} \big) \in \mathbb{R}^{n_{\text{prev}}};
\end{equation*}
similarly, the previous layer's activation values are written as the vector
\begin{equation*}
    \a_{\text{prev}} = \Big( a_{1}^{(\text{prev})}, a_{2}^{(\text{prev})}, \ldots, a_{n_{\text{prev}}}^{(\text{prev})} \Big) \in \mathbb{R}^{n_{\text{prev}}}.
\end{equation*}
In addition to its $ n_{\text{prev}} $ weights, the current neuron is parameterized by a single scalar bias $ b $, which serves as an additive constant which can translate the neuron's activation value.

@@

\begin{wrapfigure}{R}{0.36\textwidth}
    \centering
    \includegraphics[width=\linewidth]{vector/relu.pdf}
    \caption{The ReLU function}
    \label{fig:relu}
\end{wrapfigure}
Crucially, using \textit{non-linear} activation functions allows the network to form non-linear decision boundaries in the high-dimensional feature space of the inputted feature vectors; these non-linear boundaries enormously improve classification potential.
In practice, deep networks predominantly use the rectified linear unit (ReLU) function (Figure~\ref{fig:relu}) or one of its variants;\footnote{See Wikipedia's \href{https://en.wikipedia.org/wiki/Activation_function\#Comparison_of_activation_functions}{\underline{list of common activation functions}} for a comprehensive list.} I will use the notation $ f_{\text{a}} $ for generality.

@@

A discrete convolution involves two multidimensional objects---an input image and a \textit{kernel} or \textit{filter}.
The kernel is a pixel-like grid like the input image, but always smaller in height and width.
Each kernel pixel is assigned a scalar \textit{weight}, and the kernel as a whole a single scalar \textit{bias}; these kernel weights and biases are a CNN's tunable parameters, and replace the neuron weights and biases in a FCN~\cite{homl, goodfellow}.
